% !TeX spellcheck = fr_FR
% Fichier hdr.tex
% Auteur : Luiz Angelo Steffenel
% CReSTIC SysCom, Université de Reims, France
% Date de création : 17 janvier 2014
% Dernière mise à jour : 17 janvier 2014

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

\documentclass[final,twoside]{hdr} % final
%\documentclass[preliminary,nodate,oneside]{hdr}
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english,frenchb]{babel}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{subfigure}
\lstset{language=c,frame=LBtr,extendedchars=true}
%\usepackage{prog}
\usepackage{url}
%\usepackage{isolatin1}


\usepackage{amsmath}
\usepackage{amsfonts,amssymb}
\usepackage[boxed]{algorithm}
\usepackage{algorithmic}
\usepackage[]{graphics}
\usepackage{rotating}
\usepackage{multirow}

\usepackage{fancybox}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% redefinition des commandes du package algorithm & algorithmic

\renewcommand{\algorithmicif}{{\bf SI}}
\renewcommand{\algorithmicthen}{{\bf ALORS}}
\renewcommand{\algorithmicelse}{{\bf SINON}}
\renewcommand{\algorithmicend}{{\bf FIN}}
\renewcommand{\algorithmicwhile}{{\bf TANTQUE}}
\renewcommand{\algorithmicfor}{{\bf POUR}}
\renewcommand{\algorithmicforall}{{\bf POUR TOUT}}
\renewcommand{\algorithmicdo}{{\bf FAIRE}}
\renewcommand{\algorithmicelsif}{{\bf SINON SI}}

\floatname{algorithm}{Algorithme}
\renewcommand{\listalgorithmname}{Liste des algorithmes}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\newtheorem{definition}{Définition}
\newenvironment{definition}
  {\vspace{.5em} \noindent {\bf Définition : }}
  {\vspace{.5em}}

\newenvironment{remarque}
  {\vspace{.5em} \noindent {\bf Remarque : }}
  {\vspace{.5em}}

\newenvironment{lemme}
  {\vspace{.5em} \noindent {\bf Lemme : }}
  {\vspace{.5em}}

\newtheorem{corollaire}{Corollaire}
%\newtheorem{lemme}{Lemme}
\newtheorem{theoreme}{Théorème}
\newtheorem{proposition}{Proposition}
\newtheorem{propriete}{Propriété}
%\newtheorem{remarque}{Remarque}
\newcommand{\thethm}{} 

\newenvironment{resume}{%
  \noindent\rule{\linewidth}{.1pt}\par\nobreak\vskip-1em%
  \vspace{1em}\noindent\textbf{Résumé}\par\nobreak\vskip1em%
  \noindent\itshape}%
 {\par\nobreak\noindent\rule{\linewidth}{.1pt}}

%\pagestyle{myheadings}
\renewcommand{\soutname}{Soutenue publiquement le }


%\let\urlorig\url
%\renewcommand{\url}[1]{
%	\begin{otherlanguage}{english}\urlorig{#1}\end{otherlanguage}
%}

\def\imagetop#1{\vtop{\null\hbox{#1}}}

\usepackage{pstricks}
\usepackage{colortab}
\usepackage{verbatim}
\usepackage{verbatimbox}

\begin{document}

\title{Contributions à la Gestion de l'Hétérogénéité dans les Environnements Distribués \\et Pervasifs}
\author{Luiz Angelo STEFFENEL}
\date{8 décembre 2017}

\jury{
     {\em Président du jury} :                
                                        & Carine \fsc{Souveyet} &
                                        Professeur à l'Université Paris 1 Panthéon-Sorbonne\\
{\em Rapporteurs} :          &  Christophe \fsc{Cérin} &
                                        Professeur à l'Université Paris 13 - IUT Villetaneuse\\
                                        &  Emmanuel \fsc{Jeannot} & 
										Directeur de Recherche, INRIA Bordeaux Sud-Ouest\\
                                        & Philippe \fsc{Roose}   &
										Maître de Conférences HDR à l'Université de Bayonne\\
{\em Examinateurs}:         & Massimo \fsc{Villari} &
                                        Professeur à l'Università degli Studi di Messina, Italie\\
                                         & Michaël \fsc{Krajecki}   &
                                        Professeur à l'Université de Reims Champagne-Ardenne\\
	{\em Directeur} :           & Olivier \fsc{Flauzac}        & 
                                        Professeur à l'Université de Reims Champagne-Ardenne\\
                                        }
\maketitle

\begin{dedication}
à Manuele.
\end{dedication}

\begin{thanks}
{Je tiens à exprimer mes remerciements et toute ma gratitude à :}

%\item Franck Cappelo, qui m'a fait l'honneur de bien vouloir présider
%  le jury.
%
%\item P. Felber, H. Guyennet et I. Lavallée, qui ont tous trois accepté de
%  rapporter cette habilitation.
%
%\item Merci à V. Villain qui après m'avoir << initié >> à la recherche
%  il y a quelques années, a accepté de particier à mon jury.
%
%\item Je remercie A. Bui, directeur de l'équipe de recherche LICA. Son
%  soutien sans faille depuis mon arrivée à Reims, ainsi que les
%  conseils avisés qu'il m'a prodigués ont contribué à la préparation de
%  cette habilitation, ainsi qu'au développement de différents projets.
%
%\item Merci à l'ensemble de l'équipe LICA, et plus particulièrement à
%  M. Krajecki et P.-P. Mérel, avec lesquels << l'aventure >> CONFIIT a
%  été initiée et se poursuit, ainsi qu'à H.~Fouchal.
%
\item Je remercie les étudiants qui m'ont fait confiance pour
%  l'encadrement de leurs travaux de doctorat : T.~Bernard et C.~Rabat.
%
%\item Pour finir ces remerciements, je n'oublie pas ma famille, mon
%  épouse Hélène dont le soutien au quotidien à contribué à l'ensemble
%  du travail réalisé, à Michelle dont les premiers pas et les joyeux
%  << gazouillis >> ont rythmé la rédaction de cette habilitation, sans
%  oublier mon père qui a toujours cru en moi et m'a toujours soutenu,
%  et ma mère, pour laquelle j'ai une pensée particulière.
\end{thanks}

\tableofcontents

\Bigchapter{Introduction}

\input{introduction}

\chapter{L'Hétérogénéité des Communications}

\input{reseaux}

\chapter{L'Hétérogénéité des Tâches de Calcul}

\input{amide}

\chapter{L'Adaptation à la Dynamicité des Ressources}

\input{context}

\chapter{L'Hétérogénéité des Données}

\input{grappes}


\chapter{CloudFIT : un intergiciel pour le \textit{Fog Computing} et l'Internet des Objets\label{chap:CloudFIT}}

\input{cloudfit}


\Bigchapter{Conclusion et perspectives}

\input{conclusion}

\Bigchapter{Publications personnelles}

\section*{Articles dans des revues avec comité de lecture}

\subsection*{Internationales}

Steffenel, L.A., Kirsch-Pinheiro, M., Vaz Peres, L., Kirsch Pinheiro, D. "Strategies to implement Edge Computing in a P2P Pervasive Grid", International Journal of Information Technologies and Systems Approach (IJITSA), IGI Global, (accepted, to be published). 

Cassales, G. W., Charao, A., Kirsch-Pinheiro, M., Souveyet, C., Steffenel, L.A. "Improving the Performance of Apache Hadoop on Pervasive Environments through Context-Aware Scheduling ", Journal of Ambient Intelligence and Humanized Computing, Springer, 7(3), pp. 333-345, 2016. doi:10.1007/s12652-016-0361-8. 

Engel, T.A., Charao, A., Kirsch-Pinheiro, M., Steffenel, L.A. "Performance Improvement of Data Mining in Weka through Multi-core and GPU Acceleration: opportunities and pitfalls", Journal of Ambient Intelligence and Humanized Computing, Springer, June 2015. 

Diallo, T. A., Flauzac, O., Steffenel, L.A., N’diaye, S., and Dieng, Y. "GRAPP\&S, a Peer-to-Peer Middleware for Interlinking and Sharing Educational Resources". Transactions on Industrial Networks and Intelligent Systems, EAI, 2(3):e1, May 2015. 

Vasseur, R., Baud, S., Steffenel, L. A., Vigouroux, X., Martiny, L., Krajecki, M., Dauchez, M. "Inverse Docking Method for New Proteins Targets Identification: A Parallel Approach". Journal of Parallel Computing - special issue on Parallelism in Bionformatics, Elsevier, Vol 42,, pp 48-59. February 2015. 

Steffenel, L. A., Flauzac, O., Charao, A. S., P. Barcelos, P., Stein, B., Cassales, G., Nesmachnow, S., Rey, J., Cogorno, M., Kirsch-Pinheiro, M. and Souveyet, C., "Mapreduce challenges on pervasive grids", Journal of Computer Science, vol. 10 n. 11, pp. 2194-2210, July 2014. 

Vasseur, R., Baud, S., Steffenel, L. A., Vigouroux, X., Martiny, L., Krajecki, M., Dauchez, M.  "AMIDE – Automatic Molecular Inverse Docking Engine for Large-Scale Protein Targets Identification", International Journal On Advances in Life Sciences, 6 :(3\&4), IARIA, 2014.

Flauzac, O., Krajecki, M., Steffenel, L.A. "CONFIIT: a middleware for peer-to-peer computing". Journal of Supercomputing, Springer, vol 53 n. 1, July 2010, pp. 86-102. 

Nasri, W., Steffenel, L.A., Trystram, D. "Adaptive Approaches for Efficient Parallel Algorithms on Cluster-based Systems". International Journal in Grid and Utility Computing, Inderscience, vol 1 n. 2, 2009, pp 99-108. 

Steffenel, L.A., Martinasso, M., Trystram, D. "Assessing Contention Effects of All-to-All Communications on Clusters and Grids". International Journal of Pervasive Computing and Communications - Special Issue on Towards merging Grid and Pervasive Computing, Vol. 4 n. 4, 2008, pp. 440-459. 

Steffenel, L.A., Mounié, G. "A Framework for Adaptive Collective Communications for Heterogeneous Hierarchical Computing Systems". Elsevier Journal of Computer and Systems Sciences - Special Issue on Performance Analysis and Evaluation of Parallel, Cluster, and Grid Computing Systems, vol 74 n. 6, 2008, pp. 1082-1093.

%\noindent $[1]$
%M.~Krajecki, {\bf O.~Flauzac}, C.~Jaillet, P.-P.~M\'erel, and R.~Tremblay.
%{Solving an open instance of the Langford's problem using CONFIIT : a
%  middleware for peer-to-peer computing.}
% {\em Parallel Processing Letters}, accepté
%  2005.
%
%\vspace{1em} \noindent $[2]$
%H.~Baala, {\bf O.~Flauzac}, J.~Gaber, M.~Bui, and T.A. El-Ghazawi.
%{A self-stabilizing distributed algorithm for spanning tree
%construction in wireless ad hoc networks.}
% {\em Journal of Parallel and Distributed Computing}, 63(1):97--104,
%  2003.
%
%\vspace{1em} \noindent $[3]$
%H.~Baala and {\bf O.~Flauzac}.
%{Self-stabilizing algorithm for spanning tree construction with mobile
%  agents.}
% {\em Studia Informatica Universalis}, 1(1):77--82, 2001.

\subsection*{Nationales}

Vaz Peres, L., Kirsch Pinheiro, D., Steffenel, L.A., Mendes, D., Valentin Bageston, J., Dornelles Bittencourt, G., Passáglia Schuch, A., Anabor, V., Paes Leme, N.M., Schuch, N.J. "Monitoramento de Longo Prazo e Climatologia de Campos Estratosféricos quando da Ocorrência dos Eventos de Influência do Buraco de Ozônio Antártico sobre o Sul do Brasil", Revista Brasileira de Meteorologia, SBMET/Thomson-Reuters, 2017.

Flauzac, O., Steffenel, L.A., Diallo, T.H., Niang, I., Ndiaye, S. "GRAPP\&S Data Grid : Une approche de type grille et système pair-à-pair pour le stockage de données". Revue URED (Université-Recherche-Développement), Presses Universitaires de l'Université Gaston Berger de Saint-Louis du Sénégal, 2012.

\section*{Communications avec actes et comité de sélection}

\subsection*{Internationales}

Beserra, D., Kirsch-Pinheiro, M., Steffenel, L.A.,  Moreno, E.D., "Comparing the Performance of OS-level Virtualization Tools in SoC-based Systems: The Case of I/O-bound Applications", The 22nd IEEE Symposium on Computers and Communications (ISCC 2017), Heraklion, Greece, July 3-6, 2017.

Charao, A., Hoffmann, G., Steffenel, L.A, Kirsch-Pinheiro, M., Stein, B., "Performance Evaluation of Cloud-based RDBMS through a Cloud Scripting Language", 19th International Conference on Enterprise Information Systems (ICEIS), Porto, Portugal, April 26-29, 2017.

Beserra, D., Kirsch-Pinheiro, M., Steffenel, L.A., Souveyet, C., Moreno, E.D., "Performance Evaluation of OS-level Virtualization Solutions for HPC Purposes on SoC-based Systems", IEEE International Conference on  Advanced Information Networking and Applications (AINA), Taipei, Taiwan, March 27-29, 2017.

Steffenel, L.A., Kirsch-Pinheiro, M., Kirsch-Pinheiro, D., Vaz Peres, L., "Using a Pervasive Computing Environment to Identify Secondary Effects of the Antarctic Ozone Hole", 2nd Workshop on Big Data and Data Mining Challenges on IoT and Pervasive (Big2DM) , Madrid, Spain, May 23 - 26, 2016. Procedia Computer Science, v 83, pp. 1007-1012, Elsevier. 

Steffenel, L.A., Kirsch-Pinheiro, M.,  "When the Cloud goes Pervasive: approaches for IoT PaaS on a mobiquitous world",  EAI International Conference on Cloud, Networking for IoT systems (CN4IoT 2015), Rome, Italy, October 126-27, 2015. 

Steffenel, L.A., Kirsch-Pinheiro, M.,  "CloudFIT, a PaaS platform for IoT applications over Pervasive Networks",  3rd Workshop on CLoud for IoT (CLIoT 2015), Taormina, Italy, September 15, 2015. Springer CCIS 567, pp 20-32.

Steffenel, L.A., Kirsch-Pinheiro, M.,  "Leveraging Data Intensive Applications on a Pervasive Computing Platform: the case of MapReduce",  1st Workshop on Big Data and Data Mining Challenges on IoT and Pervasive (Big2DM) , London, UK, June 2 - 5, 2015. Procedia Computer Science, vol. 52, Jun 2015, Elsevier, pp. 1034–1039.  

Cassales, G.W., Charao, A.,Kirsch-Pinheiro, M., Souveyet, C., Steffenel, L.A., "Context-Aware Scheduling for Apache Hadoop over Pervasive Environments", The 6th International Conference on Ambient Systems, Networks and Technologies (ANT 2015),  London, UK, June 2 - 5, 2015. Procedia Computer Science, vol. 52, Jun 2015, Elsevier, pp. 202–209.  

Rey, J., Cogorno, M., Nesmachnow, S. and Steffenel, L.A. "Efficient Prototyping of Fault-Tolerant Map-Reduce Applications with Docker-Hadoop". WoC: First International Workshop on Container Technologies and Container Clouds,  Tempe, AZ, USA, March 9-13, 2015.

Cassales, G.W., Charao, A.,Kirsch-Pinheiro, M., Souveyet, C., Steffenel, L.A., "Bringing Context to Apache Hadoop", 8th International Conference on Mobile Ubiquitous Computing, Systems, Services and Technologies (UBICOMM 2014), Rome, Italy, August 24 - 28, 2014. ISBN: 978-1-61208-353-7, IARIA, pp. 252-258 

Engel, T.A., Charao, A., Kirsch-Pinheiro, M., Steffenel, L.A. "Performance Improvement of Data Mining in Weka through GPU Acceleration", 5th International Conference on Ambient Systems, Networks and Technologies (ANT 2014), Hasselt, Belgium, June 2 - 5, 2014. Procedia Computer Science, vol. 32, 2014, Elsevier, pp. 93–100.  

Rey, J., Cogorno, M., Nesmachnow, S. and Steffenel, L.A. "Fast Prototyping of Map-Reduce Applications with Docker-Hadoop". Conférence d’informatique en Parallélisme, Architecture et Système (ComPAS'14), April 22-25, 2014, Neuchatel, Switzerland.

Vasseur, R., Baud, S., Steffenel, L. A., Vigouroux, X., Martiny, L., Krajecki, M., Dauchez, M. "A Framework for Inverse Virtual Screening". 6th International Conference on Bioinformatics, Biocomputational Systems and Biotechnologies (BIOTECHNO 2014), Chamonix, France, 20-24 April 2014 BEST PAPER Award

Diallo, T. A., Flauzac, O., Steffenel, L.A., N’diaye, S., and Dieng, Y. "GrAPP\&S : A Distributed Framework for E-learning Resources Sharing". Proceedings of the 5th International IEEE EAI Conference on e‐Infrastructure and e‐Services for Developing Countries (AFRICOMM 2013), Blantyre, Malawi, November 25-27 2013.  Springer LNICST v. 135, pp. 219-228. 

Steffenel, L. A., Flauzac, O., Schwertner Charao, A., Pitthan Barcelos, P., Stein, B., Nesmachnow, S., Kirsch Pinheiro, M., Diaz, D. "PER-MARE: Adaptive Deployment of MapReduce over Pervasive Grids". Proceedings of the 8th International Conference on P2P, Parallel, Grid, Cloud and Internet Computing (3PGCIC'13), Compiegne, France, October 28-30 2013. 

Vasseur, R., Baud, S., Steffenel, L. A., Vigouroux, X., Martiny, L., Krajecki, M., Dauchez, M. "Parallel Strategies for an Inverse Docking Method". PBio 2013: International Workshop on Parallelism in Bioinformatics (part of EuroMPI 2013), Madrid, Spain, 15-18 September 2013. pp. 253-258

Najar, S. Kirsch-Pinheiro, M., Souveyet, C., Steffenel, L. A. "Service Discovery Mechanisms for an Intentional Pervasive Information System". Proceedings of 19th IEEE International Conference on Web Services (ICWS 2012), Honolulu, Hawaii, 24-29 June 2012. 

Steffenel, L. A., Jaillet, C., Flauzac, O., Krajecki, M. "Impact of nodes distribution on the performance of a P2P computing middleware based on virtual rings". Proceedings of the Conferencia Latino Americana de Computación de Alto Rendimiento (CLCAR 2010), Gramado, Brazil, 25-28 August 2010.

Fkaier, H., Cérin, C., Steffenel, L. A., Jemni, M. "A new heuristic for broadcasting in clusters of clusters". Proceedings of the 5th International Conference on Grid and Pervasive Computing (GPC 2010), Hualien, Taiwan, 10-14 May 2010.

Flauzac, O., Nolot, F., Rabat, C., Steffenel, L. A. "Grid of security: a new approach of the network security". Proceedings of the 3rd International Conference on Network \& System Security (NSS 2009), Gold Coast, Autralia, 19-21 October 2009. pp. 67-72

Steffenel, L. A., Kirsch-Pinheiro, M. "Strong Consistency for Shared Objects in Pervasive Grids". Proceedings of the 5th IEEE International Conference on Wireless and Mobile Computing, Networking and Communication (WiMob'2009), Marrakesh, Morroco, 12-14 October 2009. pp. 73-78 

Fathallah, K., Nasri, W., Steffenel, L. A. "On the Evaluation of OpenMP Memory Access in Multi-core Architectures". 4th International Workshop on Automatic Performance Tuning (iWAPT 2009), Tokio, Japan, 1-2 October 2009. 

Achour, S., Nasri, W., Steffenel, L. A. "On the use of performance models for adaptive algorithm selection on heterogeneous clusters". Proceedings of the 17th Euromicro International Conference on Parallel, Distributed, and Network-Based Processing (PDP 2009), Weimar, Germany, 18-20 February 2009. 

Steffenel, L.A., Kirsch-Pinheiro, M., Bebers, Y. "Total Order Broadcast on Pervasive Systems". Proceedings of the 23rd Annual ACM Symposium on Applied Computing (SAC 2008), Fortaleza, Brazil, 16-20 March 2008. pp. 2202-2206. 

Jeannot, E., Steffenel, L.A. "Fast and Efficient Total Exchange on Two Clusters". Proceedings of the 13th International Conference on Parallel Computing (EURO-PAR 2007), Rennes, France, 28-31 August 2007. Springer. LNCS 4641, pp. 848-857. 

Steffenel, L.A., Jeannot, E. "Total Exchange Performance Prediction on Grid Environments: modeling and algorithmic issues". Towards Next Generation Grid - Proceedings of the CoreGRID Symposium, Rennes, France, 27-28 August 2007. Springer, pp. 131-140.

Steffenel, L.A., Martinasso, M. and Trystram, D. "Assessing contention effects on MPI\_Alltoall communications". GPC 07 - International Conference on Grid and Pervasive Computing, Paris, France, May 2007. LNCS 4459, Springer Verlag, pp 424-435.

Nasri, W., Steffenel, L.A. and Trystram, D. "Adaptive performance modeling on hierarchical grid computing environments". 7th IEEE International Symposium on Cluster Computing and the Grid - CCGrid 2007, Rio de Janeiro, Brazil, pp 505-512, May 2007. IEEE Society.

Steffenel, L. A. "Modeling Network Contention Effects on AlltoAll Operations". IEEE Conference on Cluster Computing (CLUSTER 2006). Barcelona, Spain, 25-28 September 2006. 

Barchet-Steffenel, L. A., Mounié, G. "Scheduling Heuristics for Efficient Broadcast Operations on Grid Environments". International Workshop on Performance Modeling, Evaluation, and Optimisation of Parallel and Distributed Systems (PMEO-PDS'06), in conjunction with IPDPS'06. Rhodes Island, Greece, 25-29 Avril 2006.

Barchet-Steffenel, L. A., Mounié, G. "Total Exchange Performance Modelling under Network Contention". Proceedings of the 6th International Conference on Parallel Processing and Applied Mathematics, LNCS vol. 3911, Springer-Verlag, pp 100-107. Poznan, Pologne. 2005. 

Barchet-Steffenel, L. A., Mounié, G. "Performance Characterisation of Intra-Cluster Collective Communications". Proceedings of the SBAC-PAD 2004 16th Symposium on Computer Architecture and High Performance Computing, Foz-do-Iguacu, Brazil, IEEE Press, pp. 254-261, 2004. 

Barchet-Steffenel, L. A., Mounié, G. "Identifying Logical Homogeneous Clusters for Efficient Wide-area Communications". Proceedings of the EuroPVM/MPI 2004 11th European PVM/MPI Users' Group Meeting (2004), Budapest, Hungary, September 2004. LNCS vol. 3241, Springer-Verlag, pp. 319-326, 2004. 

Barchet-Steffenel, L. A., Mounié, G. "Fast Tuning of Intra-Cluster Collective Communications". Proceedings of the EuroPVM/MPI 2004 11th European PVM/MPI Users' Group Meeting (2004), Budapest, Hungary, September 2004. LNCS vol. 3241, Springer-Verlag, pp. 28-35, 2004. 

Barchet-Steffenel, L. A. "iRBP, A Fault Tolerant Total Order Broadcast for Large Scale Systems". Proceedings of the 9th International Conference on Parallel Computing (EURO-PAR 2003), University Klagenfurt, Klagenfurt, Austria, August 2003. LNCS vol. 2790, Springer-Verlag, pp. 632-639, 2003. 

Barchet-Steffenel, L. A.; Jansch-Pôrto, I. "On the Evaluation of heartbeat-like Detectors". Proceedings of the IEEE 2nd Latin-American Test Workshop, Cancun, México, February 2001, pp 142-147. 

%\noindent $[4]$ A.~Bui, {\bf O.~Flauzac} and C.~Rabat.  A random walk
%topology management solution for Grid computing In {\em I2CS'05,
%Innovative Internet Community Systems}, Lecture Notes on Computer
%Science. Springer, 2005, à paraître.
%
%\vspace{1em} \noindent $[5]$ T.~Bernard, A.~Bui, and {\bf O.~Flauzac}.
%{Random distributed self-stabilizing structures maintenance.}  {In
%{\em ISSADS'04, Advanced Distributed Systems: Third International
%School and Symposium}, Lecture Notes on Computer Science, volume 3061,
%pages 231--240. Springer Verlag, 2004.}
%

\subsection*{Nationales}

Nesi, L., Koslovski, G., Charão, A., Pinheiro, D., Steffenel, L.A., "Paralelização de Cálculos Estatísticos sobre Dados de Monitoramento da Camada de Ozônio: um Estudo com GPU". Fórum de iniciação científica Escola Regional de Alto Desempenho (ERAD/RS), April 5-7, 2017, Ijui, Brazil.

Muenchen, B., Siqueira, T., Nesi, L., Koslovski, G., Charão, A., Pinheiro, D., Steffenel, L.A., "Análise de Dados Observacionais sobre a Camada de Ozônio: uma Abordagem Usando MPI e OpenMP". Fórum de iniciação científica Escola Regional de Alto Desempenho (ERAD/RS), April 5-7, 2017, Ijui, Brazil.

Steffenel, L.A., Kirsch-Pinheiro, M., "Stratégies Multi-Échelle pour les Environnements Pervasifs et l’Internet des Objets". Proceedings of 11èmes Journées Francophones Mobilité et Ubiquité (Ubimob 2016), July 5, Lorient, France.

Diallo, T.H., Flauzac, O., Steffenel, L.A., Ndiaye, S., "Routage Préfixé dans GRAPP\&S". Proceedings of Colloque Africain sur la Recherche en Informatique et en Mathématiques Appliquées (CARI'2014), October 20-23, 2014, Saint Louis, Sénégal.

Diallo, T.H., Ndiaye, S., Flauzac, O., Steffenel, L.A., "GRAPP\&S, une Architecture Multi-Échelle pour les Données et le Services". Proceedings of 9èmes Journées Francophones Mobilité et Ubiquité (Ubimob 2013), June 5-6, 2013, Nancy, France. 

Najar, S. Kirsch-Pinheiro, M., Steffenel, L. A., Souveyet, C. "Analyse des mécanismes de découverte de services avec prise en charge du contexte et de l'intention". Proceedings of 8èmes Journées Francophones Mobilité et Ubiquité (Ubimob 2012), June 4-6, 2012, Anglet, France. 

Flauzac, O., Steffenel, L.A., Diallo, T.H., Niang, I., Ndiaye, S. "GRAPP\&S Data Grid : Une approche de type grille et système pair-à-pair pour le stockage de données". Colloque National sur la Recherche en Informatique et ses Applications (CNRIA'2012), Thiès/Bambey, Senegal. April 25-27, 2012.

Steffenel, L.A., Boisson, J-C., Barberot, C., Gérard, S., Hénon, E., Jaillet, C., Flauzac, O., Krajecki, M. "Deploying a fault-tolerant computing middleware over Grid’5000: performance analysis of CONFIIT and its integration with a quantum molecular docking application". 4th Grid'5000 Spring School, Reims, France, April 18-21, 2011. 

Barchet-Steffenel, L. A., Mounié, G. "Prédiction de Performances pour les Communications Collectives". Proceedings of the 16ème Rencontre Francophone du Parallélisme (RenPar'16), Le Croisic, France, pp. 101-112, April 2005. 

Barchet-Steffenel, L. A.; Jansch-Pôrto, I. "On the Evaluation of Failure Detectors Performance". Proceedings of the IX Simpósio de Computação Tolerante a Falhas (IX SCTF), Florianópolis, Brazil, March 2001, pp 73-84.

Barchet-Steffenel, L. A. "Avaliação Prática do Desempenho dos Detectores de Defeitos" (Practical Evaluation of Failure Detectors Performance). Workshop de Teses e Dissertações do IX SCTF, Florianópolis, Brazil, March 2001. 

Barchet-Steffenel, L. A., Jansch-Pôrto, I. "Comunicação Não Confiável em Detectores de Defeitos com Falhas por Crash" (Unreliable Communication in Failure Detectors with Crash Faults). Proceedings of the II Workshop de Testes e Tolerância a Falhas, Curitiba, Brazil, July 2000. 

Barchet-Steffenel, L. A., Jansch-Pôrto, I. "Avaliação Prática de um Detector de Defeitos: teoria versus implementação" (Practical Evaluation of a Failure Detector: theory versus implementation). Proceedings of the II Workshop de Testes e Tolerância a Falhas, Curitiba, Brazil, July 2000. 

Barchet-Steffenel, L. A. "Avaliação Prática dos Detectores de Defeitos e sua Influência no Desempenho das Operações de Consenso" (Practical Evaluation of Failure Detectors and their Influence on the Consensus Operations Performance). Proceedings of the V Semana Acadêmica do PPGC-UFRGS, Porto Alegre, Brazil, July 2000. 

Barchet-Steffenel, L. A. "Estudo sobre Comunicação de Grupos para Tolerância a Falhas" (A Survey on Group Communication for Fault Tolerance). Proceedings of the IV Simpósio Nacional de Informática, Centro Universitário Franciscano, Santa Maria, Brazil, 1999. 

Barchet-Steffenel, L. A. "Csockets - classes para comunicação de rede com autenticação e criptografia". Proceedings of the V Jornada Integrada de Pesquisa da UFSM (CSocketS - Communication Classes with Authentication and Cryptography), Santa Maria, Brazil, 1998.

\section*{Ouvrage Scientifique / Livres}

Steffenel, L.A., "Communications Collectives pour les Grilles de Calcul" , Éditions Universitaires Européennes, 2010. 188 pages. ISBN 978-613-1-53126-2

\section*{Chapitres de Livres}

Najar, S., Vanrompay, Y., Kirsch-Pinheiro, M., Steffenel, L.A., Souveyey, C., "Intention Prediction Mechanism in an Intentional Pervasive Information System" in K Kolomvatsos, C Anagnostopoulos, and C Hadjiefthymiades (Eds.), 'Intelligent Technologies and Techniques for Pervasive Computing, IGI Global, pp. 251-275. Mai 2013. pp. 251-275. ISBN 978-1-4666-4038-2

Flauzac, O., Nolot, F., Rabat, C., Steffenel, L.A., "Grid of Security : a decentralized enforcement of the network security" in Manish Gupta, John Walp, and Raj Sharman (Eds.), Threats, Countermeasures and Advances in Applied Information Security, IGI Global, april 2012. pp. 426-443. ISBN 9781466609785

Cérin, C., Steffenel, L.A., Fkaier, H., "Broadcasting for Grids" in Frédéric Magoules (Eds.), Fundamentals in Grid Computing: Theory, Algorithms and Technologies, Chapman \& Hall/CRC Numerical Analysis and Scientific Computing Series, chapter 8, pp 209-236, december 2009. ISBN 9781439803677

\section*{Abstracts/Posters avec actes et comité de sélection}

\subsection*{Internationales}

Vasseur, R., Haschka, T., Verzeaux, L., Albin, J., Steffenel, L.A., Khartabil, H., Belloy, N., Baud, S., Henon, E., Krajecki, M., Martiny, L., Dauchez, M., "Deciphering molecular interactions using HPC simulations: getting new therapeutic targets", 9th Ter@tec Forum, Palaiseau, France, July 1-2, 2014.

Deleau, H., Jaillet, C., Krajecki, M. and Steffenel, L.A., "Towards the Parallel Resolution of the Langford Problem on a Cluster of GPU Devices", Sixth SIAM Workshop on Combinatorial Scientific Computing (CSC14), Lyon, France, July 21-23, 2014.

Barberot, C., Boisson, J-C., Thiriot, E., Gérard, S., Monard, G., Stefenel, L-A., Hénon, E. "Study of PDE4 Inhibitors: Quantum Mechanical Molecular Docking Deployed in a Grid using CONFIIT". 9th Triennial Congress of the World Association of Theoretical and Computational Chemists (WATOC 2011). Santiago de Compostela, Spain, 17-22 July 2011. BEST POSTER AWARD

Nasri, W., Achour, S., Steffenel, L. A. "Integrating performance models and adaptive approaches for efficient parallel algorithms". 5th International Workshop on Parallel Matrix Algorithms and Applications (PMAA'08), Neuchâtel, Switzerland, pp. 18, 20-22 June 2008. 

\subsection*{Nationales}

Vasseur, R., Baud, S., Steffenel, L.A., Vigouroux, N., Martiny, L., Krajecki, M., Dauchez, M., "'Developments for a Novel Inverse Docking Method" , Journées de la Société Française de Chémoinformatique, Nancy, France, 10 et 11 octobre 2013.

Vasseur, R., Baud, S., Steffenel, L.A., Vigouroux, N., Martiny, L., Krajecki, M., Dauchez, M., "Développements HPC pour une nouvelle méthode de docking inverse" , Journée ROMEO, Université de Reims Champagne-Ardenne, Reims, France, 26 juin 2012.

Barchet-Steffenel, L. A.; Ceretta-Nunes, R. "Implementação de uma Situação de Corrida Crítica em Java" (Implementation of a Critical Run Situation in Java), Proceedings of the II Ciclo de Palestras do Curso de Informática, UFSM, Santa Maria, Brazil, october 1997

\section*{Conférences invitées}

Latin America High Performance Computing Conference (CARLA 2017)
Keynote "Harvesting the mist: expanding HPC with the help of surrounding resources". 21 september 2017.
Buenos Aires, Argentina

CLIoT/CloudWay workshops joint panel "Migrating to Cloud and IoT Solutions: Challenges and Perspectives" , 15 septembre 2015
Moderator: Nabor Mendonça, University of Fortaleza, Brazil
Panelists:
Pooyan Jamshidi, Imperial College London, U.K., \& IC4, Ireland
Maria Fazio, University of Messina, Italy
Orazio Tormachio, University of Catania, Italy
Luiz Angelo Steffenel, Université de Reims Champagne-Ardenne, France


Réunion RGE - Réseaux Grand Est. Reims, 13 février 2014.
"PER-MARE: adaptation du paradigme MapReduce aux réseaux pervasifs".

Universidade Federal de Santa Maria (UFSM, Brésil), 30 avril 2013
"O Ensino Superior em Computação na França" (L'enseignement supérieur en Informatique en France)
Rentrée solennelle du cours en Informatique à l'UFSM

IUT Villetaneuse (Université Paris 13), mai 2011
"Structure et gestion du Centre de Calcul ROMEO"

Journée ROMEO, 18 mai 2011.
"Deploying large scale application with CONFIIT: study on the Quantum Molecular Docking problem".

Universidade Federal de Santa Maria (Brésil), juillet 2010
"RemoteLabz - Environnement virtuel pour l’apprentissage des technologies réseau"

École Supérieure de Sciences et Techniques de Tunis, mai 2009.
"CONFIIT : un système pair-à-pair pour le calcul".
"GPGPU : du calcul haute performance dans votre machine".

Réunion RGE - Réseaux Grand Est. Strasbourg, 1er juin 2006.
"Modélisation des Communications de type All-to-All sujettes à la congestion du réseau".

Réunion GridExplorer (GdX). Lyon, 21 juin 2005.
"Modélisation des Communications pour les Grilles de Calcul".

Séminaires Laboratoire ID-IMAG. Grenoble, 14 novembre 2003.
"Tolérance aux Pannes et le Protocole iRBP".


\section*{Thèses et dissertations}

Barchet-Steffenel, L.A. "LaPIe : Communication Collectives Adaptées aux Grilles de Calcul". PhD Thesis, INPG, Grenoble, France, December 2005. 

Barchet-Steffenel, L.A. "Analyzing RBP, a Total Order Broadcast Protocol for Unreliable Channels". MSc. Dissertation, Doctoral School in Communication Systems, EPFL, Lausanne, Switzerland, July 2002.

Barchet-Steffenel, L. A. "Avaliação dos Detectores de Defeitos e sua Influência nas Operações de Consenso" (Evaluation of Failure Detectors and their Influence on the Consensus Operations), Master Thesis, PPGC:UFRGS, Porto Alegre, Brazil, March 2001.

Barchet-Steffenel, L. A. "CsocketS - Classes para Comunicação de Rede com Autenticação e Criptografia usando Java e CORBA" (CSocketS - Classes for Network Communication with Authentication and Criptography, using Java and CORBA). Bsc. Dissertation, UFSM, Santa Maria, Brazil, March 1999.

%\noindent $[22]$
%{\bf O.~Flauzac}
% {\em Conception d'algorithmes distribu\'es de routage tol\'erants aux
%  fautes}.
% PhD thesis, Université de Technologie de Compi\`egne,
%Janvier 2000.

\section*{Rapports de recherche}

Steffenel, L. A. "D2.1 - First Steps on the Development of a P2P Middleware for Map-Reduce". Deliverable for project STIC-AmSud PER-MARE (13STIC-07), July 2013. 

Steffenel, L. A. "Modeling Network Contention Effects on AlltoAll Operations". Rapport de recherche INRIA RR-6038 (extended version of the Cluster06 paper), November 2006. 

Steffenel, L. A. "Fast and Scalable Total Order Broadcast for Wide-area Networks". Rapport de recherche INRIA RR-6037, November 2006. 

Steffenel, L. A. "A Framework for Adaptive Collective Communications on Heterogeneous Hierarchical Networks". Rapport de recherche INRIA RR-6036 (extended version of the IPDPS06 paper), November 2006.

F. Cappello, P. Owezarski, R. Namyst, O. Richard, P. Vicat-Blanc-Primet, E. Jeannot, L.A. Steffenel, D. Caromel, P. Sens, P. Fraigniaud, C. Cérin, S. Petiton, J. Gustedt, C. Blanchet, C. Randriamaro, S. Tixieuil. "Data Grid Explorer". Rapport LAAS No05491, Projet ACI Masse de Données. Data Grid eXplorer, Septembre 2005, 48p. 

Steffenel, L. A. "Detectores de Defeitos não Confiáveis" (Unreliable Failure Detectors). Research Report (Trabalho Individual), PPGC-UFRGS, January 2000. 




\bibliographystyle{plain}
\bibliography{/Users/steffenel/Documents/Angelo-bibdesk}


\end{document}

